Title here

### 2020-07-07 analysis/paper/paper.Rmd
paper figures

---
title: "SAM simétrico y asimétrico"
author: "Elio Campitelli"
output: 
   # pdf_document:
   #    latex_engine: xelatex
   powerpoint_presentation:
    # ioslides_presentation:
        fig_height: 7.6
        fig_width: 12.8
        reference_doc: "../templates/template.pptx" 
        slide_level: 2
urlcolor: blue
header_includes:
   - \usepackage[utf8]{inputenc}
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->
<!-- 
Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`
-->

<!-- The following code chunk defines some general settings how code chunks should behave. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

@@

theme_set(theme_asymsam())
ZeroBreaks <- AnchorBreaks(0, NULL, 0)

lev.lab <- AddSuffix(" hPa")

@@

@@

@@

```{r method, fig.cap = "Spatial patterns of the first EOF of 700 hPa geopotential height. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Arbitrary units."}
sam_sep <- SAM[, eof[[1]]$right, by = .(lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev)]

@@

sam_sep %>% 
  .[lev == 700] %>%
  rm_singleton() %>% 
  melt(id.vars = c("lon", "lat")) %>%

  # .[variable == "asym", value := value + rnorm(.N, 0, sd = 1e-6)] %>%
  .[, variable := factor(variable, levels = names(lab_sam), ordered = TRUE)] %>% 
  periodic(lon = c(0, 360)) %>% 
  ggplot(aes(lon, lat)) +
  # geom_contour2(aes(z = value, linetype = factor(-sign(..level..))), global.breaks = FALSE) +
  geom_contour_fill(aes(z = value), global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = value), global.breaks = FALSE) +
  geom_qmap() +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(guide = "none") +
  coord_polar() +
  facet_grid(~variable, labeller = labeller(variable = lab_sam, lev = lev.lab)) 
```

@@

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM."}
indexes %>% 
  .[lev %in% c(30, 700) & term != "full"] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  # geom_smooth(method = "lm", aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL) +
  facet_grid(lev~., labeller = labeller(lev = lev.lab))
```

```{r asymsam-timeseries-seasonally, fig.cap = "Seasonal time series for the asymmetric SAM and symmetric SAM."}
indexes %>% 
  .[lev %in% c(30, 700) & term != "full"] %>% 
  .[, .(estimate_norm = mean(estimate_norm)), by = .(term, lev, time = seasonally(time))] %>% 
  .[, season := season(time)] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  # geom_smooth(method = "lm", aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL) +
  facet_grid(season~lev, labeller = labeller(lev = lev.lab))
```

```{r cor-lev, fig.cap = "Correlation between the Symmetric and Asymmetric SAM at each level."}
indexes_wide[, cor(asym, sym), by = lev] %>% 
  ggplot(aes(lev, V1)) +
  geom_line() +
  scale_x_level() + 
  scale_y_continuous("Correlation", limits = c(0, 1)) + 
  coord_flip()
```

```{r cross-correlation, fig.cap = "Cross correlation between levels of the Full, Asymmetric and Symmetric SAM."}
indexes %>% 
  copy() %>% 
  .[, widyr::pairwise_cor(.SD, lev, time, estimate), by = .(term)] %>% 
  ggplot(aes(item1, item2)) +
  geom_contour_fill(aes(z = correlation), na.fill = 1) +
  scale_fill_divergent("Cross-correlation") +
  scale_x_level() +
  scale_y_level() +
  coord_equal() +
  facet_grid(. ~ term, labeller = labeller(term = lab_sam))
```

```{r compute-regr-pattern}
indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev)] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)] 
```

```{r 2d-regr, fig.cap = "Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef."}
indexes_2d %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ lev, labeller = labeller(term = lab_sam, lev = lev.lab))
```

```{r compute-regr-pattern-season}
indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev, season(time))] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)]

@@

```{r 2d-regr-700, fig.cap = "Seasonal regression patterns of geopotential height at 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef."}
indexes_2d %>% 
  .[lat <= -20] %>% 
  .[lev == 700] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab))
```

```{r 2d-regr-30, fig.cap = "Seasonal regression patterns of geopotential height at 30 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef."}
indexes_2d %>% 
  .[lat <= -20] %>% 
  .[lev == 30] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20)) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-80, 80),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab))
```

```{r wave-amplitude-700, fig.cap = "Planteray wave squared amplitudes for the regression patterns at 700 hPa."}
indexes_2d %>% 
  .[ lev == 700 & lat < -20] %>%
  .[, FitWave(estimate, 1:3), by = .(lat, lev, season, term)] %>% 
  ggplot(aes(lat, amplitude^2)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Squared amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~season, labeller = labeller(term = lab_sam)) 
```

```{r wave-amplitude-30, fig.cap = "Planteray wave squared amplitudes for the regression patterns at 30 hPa."}
indexes_2d %>% 
  .[ lev == 30 & lat < -20] %>%
  .[, FitWave(estimate, 1:3), by = .(lat, lev, season, term)] %>% 
  ggplot(aes(lat, amplitude^2)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Squared amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~season, labeller = labeller(term = lab_sam)) 
```

>>>>

---
twocol: TRUE  #  Use FALSE for submission. 
title: Title here
author1: 
  name: Elio Campitelli
  email: elio.campitelli@cima.fcen.uba.ar
author2: 
  name: Leandro Díaz
affiliation: "CIMA UBA blablabla"
extra_authors: 
  - name: Carolina Vera
abstract: |
  Enter the text of your abstract here. This is a sample American Meteorological Society (AMS) \LaTeX\ template.  This document provides authors with instructions on the use of the AMS \LaTeX\ template.  Authors should refer to the file amspaper.tex to review the actual \LaTeX\ code used to create this document. The template.tex file should be modified by authors for their own manuscript.
statement: |
  This is significant becasue... 
bibliography: references.bib
output: 
  rticles::ams_article:
    latex_engine: "xelatex"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "",     # AMS required
                      out.extra = "",    # To force figure labels
                      echo = FALSE)      # Don't show R code

# This helps to use appendix figures.
appendcaption <- function (x, options) {
  `%n%` <- knitr:::`%n%`
  rw = options$resize.width
  rh = options$resize.height
  resize1 = resize2 = ""
  if (!is.null(rw) || !is.null(rh)) {
    resize1 = sprintf("\\resizebox{%s}{%s}{", rw %n% "!", 
                      rh %n% "!")
    resize2 = "} "
  }
  tikz = knitr:::is_tikz_dev(options)
  a = options$fig.align
  fig.cur = options$fig.cur %n% 1L
  fig.num = options$fig.num %n% 1L
  animate = options$fig.show == "animate"
  fig.ncol = options$fig.ncol %n% fig.num
  if (is.null(fig.sep <- options$fig.sep)) {
    fig.sep = character(fig.num)
    if (fig.ncol < fig.num) 
      fig.sep[seq(fig.ncol, fig.num - 1L, fig.ncol)] = "\\newline"
  }
  sep.cur = NULL
  if (!tikz && animate && fig.cur < fig.num) 
    return("")
  usesub = length(subcap <- options$fig.subcap) && fig.num > 
    1
  ai = options$fig.show != "hold"
  plot1 = ai || fig.cur <= 1L
  plot2 = ai || fig.cur == fig.num
  align1 = if (plot1) 
    switch(a, left = "\n\n", center = "\n\n{\\centering ", 
           right = "\n\n\\hfill{}", "\n")
  align2 = if (plot2) 
    switch(a, left = "\\hfill{}\n\n", center = "\n\n}\n\n", 
           right = "\n\n", "")
  cap = options$fig.cap
  appendix = !is.null(options$appendix)
  scap = options$fig.scap
  fig1 = fig2 = ""
  mcap = fig.num > 1L && options$fig.show == "asis" && !length(subcap)
  sub1 = sub2 = ""
  if (length(cap) && !is.na(cap)) {
    lab = paste0(options$fig.lp, options$label)
    if (plot1) {
      pos = options$fig.pos
      if (pos != "" && !grepl("^[[{]", pos)) 
        pos = sprintf("[%s]", pos)
      fig1 = sprintf("\\begin{%s}%s", options$fig.env, 
                     pos)
    }
    if (usesub) {
      sub1 = sprintf("\\subfloat[%s%s]{", subcap, knitr:::create_label(lab, 
                                                               fig.cur, latex = TRUE))
      sub2 = "}"
      sep.cur = fig.sep[fig.cur]
      if (is.na(sep.cur)) 
        sep.cur = NULL
    }
    if (plot2) {
      if (is.null(scap) && !grepl("[{].*?[:.;].*?[}]", 
                                  cap)) {
        scap = strsplit(cap, "[:.;]( |\\\\|$)")[[1L]][1L]
      }
      scap = if (is.null(scap) || is.na(scap)) 
        ""
      else sprintf("[%s]", scap)
      if (cap == "") {
        cap = ""
      } else if (appendix == TRUE) {
        cap = sprintf("\\appendcaption{%s}{%s}%s\n", options$label, cap, knitr:::create_label(lab, 
                                                                  if (mcap) 
                                                                    fig.cur, latex = TRUE))
      } else {
        cap = sprintf("\\caption%s{%s}%s\n", scap, cap, knitr:::create_label(lab, 
                                                                  if (mcap) 
                                                                    fig.cur, latex = TRUE))
      }
      fig2 = sprintf("%s\\end{%s}\n", cap, options$fig.env)
    }
  }
  else if (knitr:::pandoc_to(c("latex", "beamer"))) {
    align.env = switch(a, left = "flushleft", center = "center", 
                       right = "flushright")
    align1 = if (plot1) 
      if (a == "default") 
        "\n"
    else sprintf("\n\n\\begin{%s}", align.env)
    align2 = if (plot2) 
      if (a == "default") 
        ""
    else sprintf("\\end{%s}\n\n", align.env)
  }
  ow = options$out.width
  if (animate && identical(ow, "\\maxwidth")) 
    ow = NULL
  if (is.numeric(ow)) 
    ow = paste0(ow, "px")
  size = paste(c(sprintf("width=%s", ow), sprintf("height=%s", 
                                                  options$out.height), options$out.extra), collapse = ",")
  paste0(fig1, align1, sub1, resize1, if (tikz) {
    sprintf("\\input{%s}", x)
  }
  else if (animate) {
    aniopts = options$aniopts
    aniopts = if (is.na(aniopts)) 
      NULL
    else gsub(";", ",", aniopts)
    size = paste(c(size, sprintf("%s", aniopts)), collapse = ",")
    if (nzchar(size)) 
      size = sprintf("[%s]", size)
    sprintf("\\animategraphics%s{%s}{%s}{%s}{%s}", size, 
            1/options$interval, sub(sprintf("%d$", fig.num), 
                                    "", xfun::sans_ext(x)), 1L, fig.num)
  }
  else {
    if (nzchar(size)) 
      size = sprintf("[%s]", size)
    res = sprintf("\\includegraphics%s{%s} ", size, if (getOption("knitr.include_graphics.ext", 
                                                                  FALSE)) 
      x
      else xfun::sans_ext(x))
    lnk = options$fig.link
    if (is.null(lnk) || is.na(lnk)) 
      res
    else sprintf("\\href{%s}{%s}", lnk, res)
  }, resize2, sub2, sep.cur, align2, fig2)
}

knitr::knit_hooks$set(plot = appendcaption)
```

```{r, warning=FALSE, message = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "",
  cache.extra = 2,
  dpi = 300
)

@@

theme_set(theme_asymsam(base_size = 9))
ZeroBreaks <- AnchorBreaks(0, NULL, 0)
axis_labs_smol <- theme(axis.text = element_text(size = 7))
lev.lab <- AddSuffix(" hPa")

@@

knitr::opts_hooks$set(invisible = function(options) {
  options$include = FALSE
  options
})

@@

blabalblba introdocction. SAM... yada yada circulation.. 
`r paste(stringi::stri_rand_lipsum(3), collapse = "\n\n")`

@@

### Data

We used monthly geopotential height at 2.5 longitude by 2.5 latitude resolution from ERA5 [@hersbach].

Monthly temperature NOAA Global Surface Temperature (NOAAGlobalTemp) 5.0 degree latitude x 5.0 degree longitude global grid [@vose2012; @smith2008].

We used monthly precipitation data from CPC Merged Analysis of Precipitation [@xie1997] 2.5 degree latitude x 2.5 degree longitude.

### Definition of indexes

`r paste(stringi::stri_rand_lipsum(4, FALSE), collapse = "\n\n")`

```{r method, fig.cap = "Spatial patterns of the first EOF of 700 hPa geopotential height. Full field (left), zonally asymmetric component (middle) and zonally symmetric component (right). Arbitrary units.", fig.width=6, fig.height=3, fig.env="figure*"}
sam_sep <- SAM[, eof[[1]]$right, by = .(lev)] %>% 
  setnames("hgt", "full") %>% 
  .[, c("sym", "asym") := list(mean(full), Anomaly(full)), by = .(lat, lev)]

@@

sam_sep %>% 
  .[lev == 700] %>%
  rm_singleton() %>% 
  melt(id.vars = c("lon", "lat")) %>% 
  .[variable == "sym", value := value + rnorm(.N, 0, sd = 1e-6)] %>%
  .[, variable := factor(variable, levels = names(lab_sam), ordered = TRUE)] %>% 
  periodic(lon = c(0, 360)) %>%
  ggplot(aes(lon, lat)) +
  # geom_contour(aes(z = value, linetype = factor(-sign(..level..))), global.breaks = FALSE) +
  geom_contour_fill(aes(z = value), global.breaks = FALSE, breaks = ZeroBreaks) +
  geom_contour_tanaka(aes(z = value), global.breaks = FALSE, breaks = ZeroBreaks) +
  geom_qmap() +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(guide = "none") +
  coord_polar() +
  facet_grid(~variable, labeller = labeller(variable = lab_sam, lev = lev.lab)) +
  axis_labs_smol
```

@@

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

```{r asymsam-timeseries, fig.cap = "Time series for the asymmetric SAM and symmetric SAM.", fig.width=6, fig.height=3, fig.env="figure*"}
indexes %>% 
  .[lev %in% c(30, 700) & term != "full"] %>% 
  # .[, estimate_norm := estimate_norm/sd(estimate_norm), by = .(lev, term)] %>% 
  ggplot(aes(time, estimate_norm)) +
  geom_line(aes(color = term)) +
  scale_color_brewer(NULL, palette = "Set1", labels = lab_sam) +
  scale_y_continuous(NULL) +
  scale_x_datetime(NULL) +
  facet_grid(lev~., labeller = labeller(lev = lev.lab), scales = "free")
```

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

`r paste(stringi::stri_rand_lipsum(1, FALSE), collapse = "\n\n")`

```{r cor-lev, fig.cap = "Correlation between the Symmetric and Asymmetric SAM at each level.", fig.width=3, fig.height=3.5}
indexes_wide[, cor(asym, sym), by = lev] %>% 
  ggplot(aes(lev, V1)) +
  geom_line() +
  scale_x_level() + 
  scale_y_continuous("Correlation", limits = c(0, 1)) + 
  coord_flip()
```

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

```{r cross-correlation, fig.cap = "Cross correlation between levels of the Full, Asymmetric and Symmetric SAM.", fig.width = 6, fig.height=4, fig.env="figure*"}
tri <- matrix(c(1000, 1000, 1000, 1, 1, 1, 1000, 1000), ncol = 2, byrow = TRUE) %>% 
  as.data.table() %>% 
  setnames(., colnames(.), c("item1", "item2"))

logbreaks <- function(range) {
  ticks <- ggplot2:::calc_logticks(minpow = log10(min(range)), maxpow = log10(max(range)))
  ticks$value[ticks$end == 0.2]
}

indexes %>% 
  copy() %>% 
  .[, widyr::pairwise_cor(.SD, lev, time, estimate, diag = FALSE), by = .(term)] %>% 
  ggplot(aes(item2, item1)) +
  geom_contour_fill(aes(z = correlation), na.fill = .99, breaks = MakeBreaks(binwidth = 0.1)) +
  geom_polygon(data = tri, fill = "white") +
  scale_fill_divergent("Cross-correlation", 
                       breaks = MakeBreaks(binwidth = 0.1),
                       guide = guide_colorstrip_bottom()) +
  scale_x_level(minor_breaks = logbreaks, position = "top") +
  scale_y_level(minor_breaks = logbreaks) +
  coord_equal() +
  facet_grid(. ~ term, labeller = labeller(term = lab_sam)) +
  theme(panel.spacing.x = unit(2, "lines"), strip.placement = "outside")
```

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

```{r compute-regr-pattern}
indexes_2d_yearly <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev)] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)] 
```

```{r 2d-regr, fig.cap = "Regression patterns of geopotential height at 30, 300 and 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef.", fig.env = "figure*", fig.width=6, fig.height=7}
indexes_2d_yearly %>% 
  .[lat <= -20] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), breaks = ZeroBreaks, global.breaks = FALSE) +
  # geom_contour2(aes(z = estimate, linetype = factor(-sign(..level..))), 
  # size = 0.3,
  # global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = ZeroBreaks,
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  # geom_contour2(data = ~.x[dataset == "NCEP"], aes(z = V1)) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(), 
                       limits = c(-85, 85),
                       oob = scales::squish,
                       breaks = MakeBreaks(binwidth = 20, exclude = 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ lev, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  axis_labs_smol 
```
`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

```{r compute-regr-pattern-season}
indexes_2d <- hgt[lev %in% main_levs] %>% 
  indexes_wide[., on = .NATURAL, allow.cartesian = TRUE] %>% 
  .[, rbind(as.data.table(FitLm(hgt_a, sym, asym)),
            as.data.table(FitLm(hgt_a, full))),
    by = .(lon, lat, lev, season(time))] %>% 
  rm_intercept() %>% 
  periodic(lon = c(0, 360)) %>% 
  .[, term := factor(term, levels = names(lab_sam), ordered = TRUE)] 
```

```{r 2d-regr-700, fig.cap = "Seasonal regression patterns of geopotential height at 700 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef.", fig.width =10, fig.height=7, fig.env = "sidewaysfigure"}
indexes_2d %>% 
  .[lat <= -20] %>% 
  .[lev == 700] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate),  breaks = AnchorBreaks(0, binwidth = 5, 0),
                    global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 5, 0),
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(),
                       limits = c(-55, 55),
                       oob = scales::squish,
                       breaks = AnchorBreaks(0, binwidth = 5, 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  axis_labs_smol
```

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

```{r wave-amplitude-700, fig.cap = "Planteray wave amplitude for the regression patterns at 700 hPa.", fig.width=6, fig.height=4}
indexes_2d %>% 
  .[ lev == 700 & lat < -20] %>%
  .[, FitWave(estimate, 0:3), by = .(lat, lev, season, term)] %>% 
  ggplot(aes(lat, amplitude)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~season, labeller = labeller(term = lab_sam), scales = "free_x")
```

`r paste(stringi::stri_rand_lipsum(1, FALSE), collapse = "\n\n")`

```{r 2d-regr-30, fig.cap = "Seasonal regression patterns of geopotential height at 30 hPa with the Full, Asymmetric and Symmetric SAM. The regression patterns for Asymmetric and Symmetric SAM are the result of one multiple regression using both indices, not of two simple regressions involving each index by itsef.", fig.width =10, fig.height=7, fig.env = "sidewaysfigure"}
indexes_2d %>% 
  .[lat <= -20] %>% 
  .[lev == 30] %>% 
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate),  breaks = AnchorBreaks(0, binwidth = 20, 0),
                    global.breaks = FALSE) +
  geom_contour_tanaka(aes(z = estimate),
                      breaks = AnchorBreaks(0, binwidth = 20, 0),
                      size = 0.3,
                      global.breaks = FALSE) +
  geom_qmap(~.x[lat <= -20]) +
  scale_x_longitude() +
  scale_y_latitude(limits = c(NA, -20), labels = NULL) +
  scale_fill_divergent(NULL, guide = guide_colorstrip_bottom(),
                       limits = c(-105, 105),
                       oob = scales::squish,
                       breaks = AnchorBreaks(0, binwidth = 20, 0)) +
  scale_linetype(guide = "none") +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  axis_labs_smol
```

`r paste(stringi::stri_rand_lipsum(2, FALSE), collapse = "\n\n")`

```{r wave-amplitude-30, fig.cap = "Planteray wave amplitude for the regression patterns at 30 hPa.", fig.width=6, fig.height=4}
indexes_2d %>% 
  .[ lev == 30 & lat < -20] %>%
  .[, FitWave(estimate, 0:3), by = .(lat, lev, season, term)] %>% 
  ggplot(aes(lat, amplitude)) +
  geom_line(aes(color = factor(k))) +
  scale_color_brewer("Wave-number", palette = "Set1") +
  scale_y_continuous("Squared amplitude") +
  scale_x_latitude(ticks = 15) +
  coord_flip() +
  facet_grid(term~season, labeller = labeller(term = lab_sam), scales = "free_x") 
```
`r paste(stringi::stri_rand_lipsum(1, FALSE), collapse = "\n\n")`

```{r compute-pp-regr} 
pp <- CMAP() %>% 
  ReadNetCDF(subset = list(lat = c(-90, -20))) %>% 
  normalise_coords() %>% 
  .[, precip := Anomaly(precip), by = .(lon, lat)]

pp_regr_season <- pp %>% 
  # .[, air := Anomaly(air), by = .(lon, lat)] %>%
  .[indexes_wide[lev == 700], on = .NATURAL] %>% 
  rm_singleton() %>% 
  .[, lapply(.SD, mean, na.rm = TRUE), by = .(lon, lat, time = seasonally(time))] %>%
  .[, nas := mean(is.na(precip)), by = .(lon, lat, season(time))] %>% 
  .[nas >= 0.15, precip := NA] %>% 
  .[, rbind(as.data.table(FitLm(precip, asym, sym, se = TRUE)),
            as.data.table(FitLm(precip, full, se = TRUE))),
    by = .(lon, lat, season(time))] %>% 
  rm_intercept() %>% 
  .[, p.value := Pvaluate(estimate, std.error, df, "fdr"), by = .(season, term)]

@@

```{r pp-regr-global, fig.cap = "caption", fig.width =10, fig.height=7, fig.env = "sidewaysfigure"}
pp_regr_season %>%
  ggplot(aes(lon, lat)) +
  geom_contour_fill(aes(z = estimate), na.fill = 0, breaks = AnchorBreaks(0, 0.25, 0)) +
  geom_contour2(aes(z = p.value), breaks = 0.01) +
  geom_contour_fill(aes(z = as.numeric(is.na(estimate))), breaks = 1) +
  stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)), size = 0.4, alpha = 0.3) +
  geom_qmap(~.x[lat <= -20]) +
  scale_x_longitude() +
  scale_y_latitude(labels = NULL) +
  scale_fill_divergent("Precipitation", guide = guide_colorstrip_bottom(), 
                       breaks = AnchorBreaks(0, 0.25, 0)) +
  coord_polar() +
  facet_grid(term ~ season, labeller = labeller(term = lab_sam, lev = lev.lab)) +
  axis_labs_smol
```

```{r regr-pp-season, fig.cap = "Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas.", invisible = TRUE}

is.sa <- function(data, lons = c(265, 325), lats = c(-80, -10)) {
  data[, lon %between% lons & lat %between% lats]
}

filter_sa <- function(data, lons = c(265, 325), lats = c(-80, -10)) {
  data[is.sa(data, lons = lons, lats = lats)]
}

pp_regr_season %>% 
  filter_sa() %>% 
  .[, lon := ConvertLongitude(lon)] %>% 
  ggplot(aes(lon, lat)) +
  # geom_raster(aes(fill = estimate)) +
  geom_contour_fill(aes(z = estimate), na.fill = 0, breaks = AnchorBreaks(0, 0.25, 0)) +
  geom_contour2(aes(z = p.value), breaks = 0.01) +
  geom_contour_fill(aes(z = as.numeric(is.na(estimate))), breaks = 1) +
  stat_subset(aes(subset = p.value <= 0.01 & is.cross(lon, lat)), size = 0.4, alpha = 0.3) +
  geom_qmap(~filter_sa(.x)[, lon := ConvertLongitude(lon)], keep = 0.02) +
  scale_x_longitude() +
  scale_y_latitude() +
  scale_fill_divergent("Precipitation", guide = guide_colorstrip_bottom(), 
                       breaks = AnchorBreaks(0, 0.25, 0)) +
  # coord_polar() +
  ggalt::coord_proj("+proj=moll +lon_0=-60") +
  facet_grid(season~term, labeller = labeller(term = lab_sam)) +
  axis_labs_smol
```

```{r compute-air-regr}
time_subset <- list(time = range(indexes$time))
air <- NOAATemp() %>% 
  ReadNetCDF(vars = "air", subset = c(time_subset, list(lat = c(-90, 10)))) %>% 
  normalise_coords()

air_regr <- air %>% 
  # .[, air := Anomaly(air), by = .(lon, lat)] %>%
  .[indexes_wide[lev == 700], on = .NATURAL] %>% 
  rm_singleton() %>% 
  .[, lapply(.SD, mean, na.rm = TRUE), by = .(lon, lat, time = seasonally(time))] %>%
  .[, nas := mean(is.na(air)), by = .(lon, lat, season(time))] %>% 
  .[nas >= 0.15, air := NA] %>% 
  .[, rbind(as.data.table(FitLm(air, asym, sym, se = TRUE)),
            as.data.table(FitLm(air, full, se = TRUE))),
    by = .(lon, lat, season(time))] %>% 
  rm_intercept() 
```
